---
title: "Class_two"
author: "Emily Ferry"
date: "1/20/2020"
output: 
md_document:
  variant: gfm
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
#install.packages("gapminder")
library(gapminder)
library(ggplot2)
library(tidyverse)
#install.packages("ggthemes")
library(ggthemes)
options(sypen=999)
?gapminder#Excerpt of the Gapminder data on life expectancy, GDP per capita, and population by country.


```
### intro to GGPlot2

## gramar of graphics 
 
He's going to push his file to github for us to reference so we dont need to copy 

*Start with a ggplot 
*add data and arstedic mapping 
*geometric objests
*scales 
*facet specs
*statistical transormation 
*coordinate system 


#Lets get the data ready 

```{r gapminder}

data("gapminder")
str(gapminder) #6 variables, 2 factors, year int, lifeexpectancey, gdp per cap, etc. 

```


```{r}
#ggplot2::  #reference the package without lib. 
ggplot(data=gapminder, mapping = aes(x = year, y = pop)) +
  geom_point()                              #put in some points 


ggplot()+
  geom_point(data = gapminder, mapping = aes( x= year, y= pop)) # this is useful for layering two datasets, you can do with smaller subsets of the data for layers 

ggplot()+
  geom_point(data = gapminder, mapping = aes( x= year, y= pop, color= continent))

 ggplot(data = gapminder, mapping = aes( x= year, y= pop, color= continent))+
   geom_point()
 
 


```


#Aggregate by contients 
```{r}

gapminder %>%
  group_by(year,continent) %>% 
    summarise(pop= sum(pop)) %>% 
   ggplot(data = ., mapping = aes( x= year, y= pop, color= continent))+
   geom_point() # our scale here is way off, not very useful to show 
# make a transformation to the Y - a log 

gapminder %>%
  group_by(year,continent) %>% 
    summarise(pop= sum(pop)) %>% 
   ggplot(data = ., mapping = aes( x= year, y= pop, color= continent))+
   geom_point() +
  scale_y_log10()  # now we can see oceian, but at the cost of other things 



view(gapminder)
summary(gapminder)

```
# looking at other variables - GDP per Cap 
```{r}
ggplot(data = gapminder, mapping = aes( x= year, y=gdpPercap, color = country))+
  geom_point() +      #facet by countires - grip = two variabels, wrap = only one 
  facet_wrap(~continent)
  
# this got a loter nutty - we want to filter by Asia 

gapminder%>%
  filter(continent == "Asia") %>% 
ggplot(data = ., mapping = aes( x= year, y=gdpPercap, color = country))+
  geom_point() +      #facet by countires - grip = two variabels, wrap = only one 
  facet_wrap(~continent)


```
#try two numeric values

```{r}
ggplot(data = gapminder, aes(x=gdpPercap, y = lifeExp)) +
  geom_point() # two numeric values - see some interesting realtionship kinda weird - well try the log 

ggplot(data = gapminder, aes(x=gdpPercap, y = lifeExp)) +
  geom_point() + 
  scale_x_log10()  # this is alot easier to read - so an example of when you'd want to do it, check it out by contient 

ggplot(data = gapminder, aes(x=gdpPercap, y = lifeExp, color = continent)) +
  geom_point() + 
  scale_x_log10() +  # this is better yet
  stat_smooth() # add the smoother line in, default is the loess smoother, local regressions, we can change that if we want 

ggplot(data = gapminder, aes(x=gdpPercap, y = lifeExp, color = continent)) +
  geom_point() + 
  scale_x_log10() +  # this is better yet
  stat_smooth( method = lm) # life expetancy on the log of gdp per cap ( using a linear model)


# you can make this graph even more confusing/ advanced 

ggplot(data = gapminder, aes(x=gdpPercap, y = lifeExp, color = continent, size = pop)) +
  geom_point() + 
  scale_x_log10() +  # this is going to be bad - we increase point size by the population weight 
  stat_smooth( method = lm)

gapminder%>%
  filter(continent == "Asia") %>% 
ggplot(data = ., mapping = aes( x= year, y=gdpPercap, color = country))+
  geom_point() +      #facet by countires - grip = two variabels, wrap = only one 
  facet_wrap(~continent)
```
# changing mappong for pop 
```{r}
gapminder %>%
  filter( continent== "Europe") %>%
  group_by(country) %>% 
  summarise(gdpPercap = mean (gdpPercap), 
lifeExp = mean( lifeExp))%>%
ggplot(data = . , aes(x=gdpPercap, y = lifeExp, shape = country)) +
  geom_point() + 
  scale_x_log10() + 
  stat_smooth( method = lm)

#Not really usefuly, but we can see that there are multiple shapes though 


```
# Back to square one 

```{r}

ggplot(data = gapminder , mapping=  aes(x=gdpPercap, y = lifeExp, fill = continent)) +
  geom_point() + 
  scale_x_log10() +
  stat_smooth( method = lm) # cool to see that colors 



ggplot(data = gapminder , mapping=  aes(x=gdpPercap, y = lifeExp, color = continent)) +
  scale_x_log10() +
  stat_smooth( method = lm) # to get ride of points omit the geom_point 

# dashed lines, hard to tell though which is which, cause it kind of makes it up 

ggplot(data = gapminder , mapping=  aes(x=gdpPercap, y = lifeExp, linetype = continent)) +
  geom_point(aes( color = continent))+   # you can add the color in the points 
  scale_x_log10() +
  stat_smooth( method = lm)

#getting rid of the lines 

ggplot(data = gapminder , mapping=  aes(x=gdpPercap, y = lifeExp)) +
  geom_point(aes( color = continent, alpha = pop))+  # looks mysterious, alpha can take a fixed value for the color or you can change he hue of each point  
  scale_x_log10() +
  stat_smooth( method = lm)


ggplot(data = gapminder , mapping=  aes(x=gdpPercap, y = lifeExp)) +
  geom_point(aes( color = continent), alpha = .5) +  # if you move the alpha outside of the aes () then you can control the weight/ hue, alpah ranges from 0-1 
  scale_x_log10() +
  stat_smooth( method = lm, mapping = aes ( color = continent))

```
```{r}
ggplot(data = gapminder , mapping=  aes(x=gdpPercap, y = lifeExp)) +
  geom_point( alpha = .4) + 
  scale_x_log10() +
  stat_smooth( method = lm) + 
  facet_wrap(~continent) # added a facet, shows the relationships, and breaks the scatter plots out 


ggplot(data = gapminder , mapping=  aes(x=gdpPercap, y = lifeExp)) +
  geom_point( alpha = .4) + 
  scale_x_log10() +
  stat_smooth( method = lm) + 
  facet_wrap(~continent)+ 
  xlab("GDP Per Capita")+ 
  ylab ("Life Expectancy")+
  ggtitle("Life Expectancy by GDP Per Capita", subtitle = "Faceted by Continet") # adding the titles, and the axisis 


```
# Getting Weird again 


```{r}
ggplot(data = gapminder , mapping=  aes(x=gdpPercap, y = lifeExp)) +
  geom_point( alpha = .4) + 
  scale_x_log10() +
  stat_smooth( method = lm) + 
  facet_wrap(~continent)+ 
  xlab("GDP Per Capita")+ 
  ylab ("Life Expectancy")+
  ggtitle("Life Expectancy by GDP Per Capita", subtitle = "Faceted by Continet")+ 
  coord_flip() #flipping the axis 

```

```{r}
gapminder %>% 
ggplot(data = . , mapping=  aes(x= continent, y = lifeExp)) +
  geom_boxplot() # here we can see the box plots, but if we flip its a little easier to read 

gapminder %>% 
ggplot(data = . , mapping=  aes(x= continent, y = lifeExp)) +
  geom_boxplot() + 
  coord_flip() # yes, this is easier to read for sure 

gapminder %>% 
ggplot(data = . , mapping=  aes(x= continent, y = lifeExp)) +
  geom_boxplot() + 
  coord_flip()+
  theme_minimal() #took out the grey background, classic would remove all of the lines


gapminder %>% 
ggplot(data = . , mapping=  aes(x= continent, y = lifeExp)) +
  geom_point()+
  geom_boxplot() + 
  coord_flip()+
  theme_minimal() # a bit much, but lets you see a little more on the data, we can change this if we wanted 

gapminder %>% 
ggplot(data = . , mapping=  aes(x= continent, y = lifeExp)) +
  geom_point()+
  geom_boxplot() + 
  coord_flip()+
  theme_fivethirtyeight()
# west andersone themes, you can have plots in the different themes and colors etc. 





```

